### 2018-09-12 main.tex

\documentclass{article}

\title{My title}
\author{Alexey N. Shiklomanov\(^1\), Elizabeth M. Cowdery\(^1\), Michael
  Bahn\(^2\), Chaeho Byun\(^3\), Joseph Craine\(^4\), Andrés
  Gonzalez-Melo\(^5\), Steven Jansen\(^6\), Nathan Kraft\(^7\), Koen
  Kramer\(^{8}\), Vanessa Minden\(^{9,10}\), Ülo Niinemets\(^{11}\),
  Yusuke Onoda\(^{12}\), Enio Egon Sosinski\(^{13}\), Nadejda A.
  Soudzilovskaia\(^{14}\), Michael C. Dietze\(^1\)}

\begin{document}

% Affiliations
\(^1\)Department of Earth \& Environment, Boston University, Boston, MA
02215, USA;
\(^2\)Institute of Ecology, University of Innsbruck, 6020
Innsbruck, Austria;
\(^3\)School of Biological Sciences, Seoul National
University, Gwanak-ro 1, 08826, Seoul, Seoul, South Korea;
\(^4\)Jonah
Ventures, Manhattan, KS 66502, USA;
\(^5\)Universidad del Rosario.
Facultad de Ciencias Naturales y Matemáticas. Carrera 26 \#63B-48.
Bogotá, Colombia;
\(^6\)Institute of Systematic Botany and Ecology, Ulm
University, Ulm, Germany;
\(^7\)Department of Ecology and Evolutionary
Biology, University of California Los Angeles, Los Angeles, CA 90095,
USA;
\(^{8}\)Vegetation, Forest, and Landscape Ecology, Wageningen
Environmental Research and Wageningen University, Wageningen,
Gelderland, The Netherlands;
\(^{9}\)Institute for Biology and
Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl
von Ossietzky Str. 9-11, 26129 Oldenburg, Germany;
\(^{10}\)Department
of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan
2, 1050 Brussels;
\(^{11}\)Institute of Agricultural and Environmental
Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014
Tartu, Estonia;
\(^{12}\)Graduate School of Agriculture, Kyoto
University, Kyoto, 606-8502, Japan;
\(^{13}\)Embrapa Temperate
Agriculture, Rodovia BR 392, Km 78, CEP 96010-971 Pelotas, RS, Brazil;
\(^{14}\)Conservation Biology Department, Institute of Environmental
Sciences, Leiden University, Leiden, The Netherlands

Corresponding author:

Alexey Shiklomanov,

telephone: +1 (302) 897-3348,

email: ashiklom@bu.edu

Key words: Functional trade-off; hierarchical modeling; trait variation;
ecological modeling; leaf morphology; leaf biochemistry

\begin{longtable}[]{@{}lll@{}}
\toprule
Section & Word count &\tabularnewline
\midrule
\endhead
Introduction & 1060 & Figures: 5 (all colored)\tabularnewline
Materials and Methods & 1695 & Tables: 2\tabularnewline
Results & 1108 & Supplementary figures: 2\tabularnewline
Discussion & 1907 & Supplementary tables: 4\tabularnewline
Acknowledgements & 88 & Supplementary methods: 1\tabularnewline
Total & 5858 &\tabularnewline
\bottomrule
\end{longtable}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: