### 2019-09-09 authors.tex

\usepackage[noblocks]{authblk}

\author[1,*]{Alexey N. Shiklomanov <alexey.shiklomanov@pnnl.gov>}
\affil[1]{Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD}

\author[2]{Elizabeth M. Cowdery <ecowdery@bu.edu>}
\affil[2]{Department of Earth \& Environment, Boston University, Boston, MA}

\author[3]{Michael Bahn <Michael.Bahn@uibk.ac.at>}
\affil[3]{Institute of Ecology, University of Innsbruck, 6020 Innsbruck, Austria}

\author[4]{Chaeho Byun <chaeho.byun@mail.mcgill.ca>}
\affil[4]{School of Civil and Environmental Engineering, Yonsei University, Seoul 03722, Korea}

\author[5]{Steven Jansen <steven.jansen@uni-ulm.de>}
\affil[5]{Institute of Systematic Botany and Ecology, Ulm University, Albert-Einstein-Allee 11, 89081, Ulm, Germany}

\author[6]{Koen Kramer <koen.kramer@wur.nl>}
\affil[6]{Department of Vegetation, Forest, and Landscape Ecology, Wageningen Environmental Research and Wageningen University, Wageningen, Gelderland, The Netherlands}

\author[7,8]{Vanessa Minden <vanessa.minden@uni-oldenburg.de>}
\affil[7]{Institute for Biology and Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl von Ossietzky Str. 9-11, 26129 Oldenburg, Germany}
\affil[8]{Department of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussels}

\author[9]{\"Ulo Niinemets <ylo.niinemets@emu.ee>}
\affil[9]{Institute of Agricultural and Environmental Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014 Tartu, Estonia}

\author[10]{Yusuke Onoda <yusuke.onoda@gmail.com>}
\affil[10]{Graduate School of Agriculture, Kyoto University, Kyoto, 605-8503, Japan}

\author[11]{Nadejda A. Soudzilovskaia <n.a.soudzilovskaia@cml.leidenuniv.nl>}
\affil[11]{Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands}

\author[2]{Michael C. Dietze <dietze@bu.edu>}

\affil[*]{Corresponding author; Phone: (301) 314-6713; Fax: (301) 314-6719; Email: alexey.shiklomanov@pnnl.gov; Mail: 5825 University Research Ct., Office 3533, College Park, MD 20740}

\renewcommand\Authfont{\footnotesize}
\renewcommand\Affilfont{\scriptsize}