### 2019-09-09 AppendixS1.tex
Fix Appendix S1 formatting.

% Created 2019-09-09 Mon 09:26
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage[backend=biber,style=authoryear,date=year]{biblatex}
\addbibresource{/Users/shik544/Dropbox/references/library.bib}
\input{authors.tex}
\date{}
\title{Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis}
\hypersetup{
 pdfauthor={},
 pdftitle={Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.3 (Org mode 9.3)}, 
 pdflang={English}}
\begin{document}

\maketitle
\noindent \textbf{Journal:} Ecological Applications

\clearpage

\noindent
\textbf{Appendix S1:} References for the original sources of the TRY trait data used in our analysis.

\vspace{\baselineskip}

\nocite{kleyer_2008_leda_traitbase}

\nocite{wirth_2009_imprint}

\nocite{baraloto_2010_decoupled}

\nocite{craine_2009_global}

\nocite{veihmeyer_1956_soil_moisture}

\nocite{cerabolini_2010_can}

\nocite{wright_2010_functional}

\nocite{wright_2004_worldwide}

\nocite{ogaya_2003_comparative}

\nocite{akhmetzhanova_2012_rediscovered_treasure}

\nocite{frenette-dussault_2011_functional}

\nocite{medlyn_1999_effects}

\nocite{kattge_2009_quantifying}

\nocite{laughlin_2011_climatic}

\nocite{campetella_2011_patterns}

\nocite{auger_2012_inter}

\nocite{fyllas_2009_basin}

\nocite{blonder_2012_leaf}

\nocite{cornelissen_2003_functional}

\nocite{cavender-bares_2006_phylogenetic}

\nocite{garnier_2007_assessing}

\nocite{prentice_2010_evidence}

\nocite{brown_2011_assessing}

\nocite{schweingruber_2005}

\nocite{onoda_2011_global}

\nocite{han_2005_leaf}

\nocite{domingues_2010_co}

\nocite{han_2005_leaf}

\nocite{messier_2010_how}

\nocite{price_2007_scaling}

\nocite{kraft_2008_functional}

\nocite{pierce_2013_allocating}

\nocite{wright_2006_relationships}

\nocite{reich_2008_scaling}

\nocite{domingues_2013_seasonal}

\nocite{niinemets_2001_global}

\nocite{cornwell_2008_plant}

\nocite{vergutz_2012_global}

\nocite{sandel_2011_using}

\nocite{poorter_2009_causes}

\nocite{gallagher_2012_global}

\nocite{williams_2012_lba}

\nocite{choat_2012_global}

\nocite{fonseca_2000_shifts}

\nocite{burrascano_2015_wild}

\nocite{milla_2011_multi}

\nocite{adriaenssens_2012_dry}

\nocite{willis_2009_phylogenetic}

\nocite{kisel_2012_testing}

\nocite{pierce_2012_combined}

\nocite{cornelissen_1996_experimental}

\nocite{yguel_2011_phytophagy}

\nocite{xu_2003_seasonal}

\nocite{bond-lamberty_2004_net}

\nocite{von_2004_testing}

\nocite{chen_2011_leaf}

\nocite{byun_2012_plant}

\nocite{laughlin_2011_climatic}

\nocite{ordonez_2010_plant}

\nocite{kerkhoff_2006_phylogenetic}

\nocite{butterfield_2010_regeneration}

\nocite{bond-lamberty_2002_aboveground}

\nocite{pierce_2007_disturbance}

\nocite{gutierrez_2012_successional}

\nocite{manzoni_2013_optimization}

\nocite{wright_2012_does}

\nocite{chambers_2004_respiration}

\nocite{guerin_2012_leaf}

\nocite{bond-lamberty_2002_leaf}

\nocite{mueller_2006_plant}

\nocite{cerabolini_2010_can}

\nocite{craine_2011_functional}

\nocite{craine_2012_community}

\nocite{shipley_2002_trade}

\nocite{wohlfahrt_1999_inter}

\nocite{rueger_2011_determinants}

\nocite{rueger_2009_response}

\nocite{quested_2003_decomposition}

\nocite{choat_2012_global}

\nocite{pahl_2013_no}

\nocite{shipley_2002_dry}

\nocite{minden_2011_testing}

\nocite{meir_2002_acclimation}

\nocite{schererlorenzen_2007_exploring}

\nocite{bakker_2006_plant}

\nocite{penuelas_2009_faster}

\nocite{powers_2010_plant}

\nocite{shiodera_2008_variation}

\nocite{fitter_1994_ecological_flora_database}

\nocite{minden_2012_plant}

\nocite{bond-lamberty_2006_nitrogen}

\nocite{rueger_2011_determinants}

\nocite{peco_2005_effect}

\nocite{tucker_2011_physiological}

\nocite{weedon_2009_global}

\nocite{blonder_2010_venation}

\nocite{spasojevic_2012_inferring}

\nocite{campbell_2007_acclimation}

\nocite{craven_2007_between}

\nocite{cornelissen_2004_leaf}

\nocite{pillar_2003_improved}

\nocite{hickler_1999_plant}

\nocite{bakker_2005_effects}

\nocite{vile_2005_significations}

\nocite{craine_2005_environmental}

\nocite{swaine_2007_ecological_bornean}

\nocite{wilson_2000_spatial}

\nocite{preston_2006_wood}

\nocite{zapata-cuartas_2012_probability}

\nocite{blonder_2013_assessing}

\nocite{blonder_2010_venation}

\nocite{guy_2013_limited}

\nocite{kichenin_2013_contrasting}

\nocite{louault_2005_plant}

\nocite{araujo_2012_lba}

\nocite{bocanegra_2015_grupos}

\nocite{beckmann_2012_role}

\nocite{adler_2004_functional}

\nocite{loveys_2003_thermal}

\nocite{bodegom_2008_separating}

\nocite{domingues_2013_seasonal}

\nocite{shipley_2000_functional}

\nocite{loranger_2012_predicting}

\nocite{pyankov_1999_leaf}

\nocite{shipley_1995_structured}

\nocite{cornelissen_1996_seedling}

\nocite{carswell_2000_photosynthetic}

\nocite{givnish_2004_adaptive}

\nocite{meziane_1999_interacting}

\nocite{meir_2007_photosynthetic}

\nocite{atkin_1999_response}

\nocite{diaz_2004_plant}

\nocite{kazakou_2006_co}

\nocite{demey_2013_nutrient}

\printbibliography[heading=none]
\end{document}